---
title: "Interview - Developmental Milestone Project"
subtitle: "How to Measure Parents' Developmental Foci: Domains and Methodologies"
author: "Chi Zhang"
institute: the University of Manchester
date: today
date-format: "dddd, [the] D[<sup style='font-size:65%;font-style:italic;'>th</sup>] [of] MMMM, YYYY"
bibliography: bib/bib.bib
csl: https://www.zotero.org/styles/apa
format:
  dime-revealjs:
    embed-resources: true
    footer: " "
    self-contained-math: false
    output-file: index.html
    toc: true
    toc-depth: 1
    toc-title: "Contents"
    preview-links: false

    html-math-method: mathjax
editor: 
  markdown: 
    wrap: 72
---

```{r Setup}
#| include: false
# Add default ggplot theme. This can be overridden by your own theme at
# the end of the ggplot graph.

# DIME:
source("_extensions/dime-worldbank/dime/setup_dime_palettes.R")
source("_extensions/dime-worldbank/dime/setup_ggplot2_dime.R")

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  dplyr, DT, ggplot2, ggpubr, ggrepel, ggtext, gt, here, huxtable, 
  knitr, leaflet, osmdata, pacman, pagedown, palmerpenguins,
  reactable, sf, tidyr, tidyverse, brms, rstan, patchwork
)
```

## About me

-   **PhD in Education**, The University of Manchester (2021–Present);\
    *From belief to practice: Understanding teachers’ pedagogic choices
    in China’s mathematics education*

-   **MSc in Research Methods with Education** (Distinction,
    2020–2021);\
    *The association between mathematics teachers’ exclusivity belief
    and pedagogic practice*

-   **MA in Mathematics and Pedagogy**, The Education University of Hong
    Kong (2019–2020)

## Interview question

1.  How can one assess parents' developmental foci in a cross-cultural
    context? With developmental foci, we mean the domain (e.g.
    linguistic, social, physical, etc.) of a child’s development that is
    of particular importance to parents.
2.  Think about two communities (could be Germany and Namibia, but also
    other contexts you have experience with) and formulate hypotheses
    regarding potential differences in developmental foci.

# The basic IRT assumptions {background-color="#07202E"}

## What is parents' development foci?

::: panel-tabset
### Developmental Niche

-   Physical and social settings of the child’s daily life;
-   Customs and practices of care; and
-   **The psychology of the caretakers [@super1986develo].**

### Parental Ethnotheories

-   Cultural models of parenting and child development held by parents
    and others [@harkness1996parent];

-   Also link to parents' cultural belief
    systems/expectations/goals/focus etc.

-   "Within the constraints imposed by the wider environment, parents
    make choices about **the best ways** to take care of their children,
    and these choices tend to follow **culturally recognizable
    patterns** [@harkness2020why, pp.18]."
:::

## The basic assumptions of cultural models

:::: panel-tabset
### Unidimensional spectrum

Individualism/collectivism [e.g., @triandis1989self]; distal/proximal
parenting [@keller2009cultur]; autonomy/relatedness [@keller2006cultur]
...

-   **+** : Easy to model `(1|person)`;

-   **-** : Too simplified, criticisms argue autonomy and relatedness
    can coexist and need to coexist because they are both human needs
    [@oyserman2002rethin; @keller2016psycho].

### Example

::: {style="font-size: 0.7em;"}
Data from [@ndzenyuiy2026parent]:

| Parental Ethnotheory Questionnaire (PEQ) Item | Factor Loading |
|:-----------------------------------------------------|:----------------:|
| It is important to rock a crying baby on the arms in order to console him/her. | 0.000 |
| Sleeping through the night should be trained as early as possible. | -0.209 |
| It is not necessary to react immediately to a crying baby. | **-0.321** |
| You cannot start early enough to direct the infant’s attention towards objects and toys. | -0.152 |
| Gymnastics make a baby strong. | **0.507** |
| If a baby is fussy, he/she should be immediately picked up. | **0.767** |
| It is good for a baby to sleep alone. | 0.000 |
| When a baby cries, he/she should be narrated immediately. | **0.551** |
| Babies should be left crying for a moment in order to see whether they console themselves. | -0.175 |
| A baby should be always in close proximity with his/her mother, so that she can react immediately to his/her signals. | **0.424** |
:::
::::

## The basic assumptions of cultural models

:::: panel-tabset
### Multidimensional construct

For example, Kağıtçıbaşı’s [-@kagitcibasi2017family] four-field scheme
with variations in the dimensions of agency (autonomy & heteronomy) and
interpersonal distance (closeness & separateness).

-   **independent self**: autonomy + separateness (western middle-class,
    *Germany*)
-   **interdependent self**: heteronomy + closeness (rural farming
    community, *Namibia*)
-   **autonomous-related self**: autonomy + closeness (non-western urban
    community, *China*)
-   **separated heteronomous self**: heteronomy + separateness

### Example

::: {style="font-size: 0.6em;"}
How important is it that your child, when an adult [@liang2021measur]...

| Item | Item (stem continuation) | Dimension |
|--------------:|:------------------------------------------|:--------------|
| 1 | … maintains good relationships with many people? | Relatedness |
| 2 | … cares about others’ feelings? | Relatedness |
| 3 | … is loyal to his or her friends? | Relatedness |
| 4 | … feels well connected to other people? | Relatedness |
| 5 | … is well connected to the extended family (grandparents, aunts, cousins, etc.)? | Relatedness |
| 6 | … tries to reach his or her goals without anyone else’s help? | Autonomy |
| 7 | … tries not to depend on someone else to achieve his or her goals? | Autonomy |
| 8 | … typically decides on a course of action without help from others? | Autonomy |
| 9 | … makes decisions about what to do without being influenced by others’ opinions? | Autonomy |
| 10 | … likes to live without many ties to others? | Separateness |
| 11 | … prefers to live alone? | Separateness |
| 12 | … keeps personal issues to himself or herself? | Separateness |
| 13 | … does things in traditional ways? | Heteronomy |
| 14 | … does the things that other people expect of him or her? | Heteronomy |
| 15 | … avoids doing things that other people say are wrong? | Heteronomy |
:::
::::

# Some concerns {background-color="#07202E" background-opacity="0.8"}

## Issues in IRT

-   Optimistic/Conservative responding style
-   Ceiling effect
-   Anchoring

## Strange Results in PISA 2012

![From [@kyllonen2015anchor]](figures/PISA.png){width="80%" fig-align="center" style="margin-top: 500px;"}

## Example of anchors

-   Anchor in category (How often do you.... 'half of the time')
-   Anchor in items [items from free listing, @vandevijver2004bias]
-   Anchor in context (We would like to know which of the following
    characteristics are important for your child when he or she has
    become an adult...; )
-   Anchor in modelling (`Comm` in fixed/random effect)

# Simulation and visualisation {background-color="#07202E"}

## Simulation Settings (N=200)

-   **Sample & Scale:**
    -   Groups: Germany (g, $n=100$) \| Namibia (n, $n=100$).
    -   Scale: 30 items (10 per domain), 4-point Likert, **RSM (acat)**
        logic.
-   **Latent Structure (Traits):**
    -   Domains: **L** (Linguistics), **S** (Social), **P**
        (Professional).
    -   Consistency: $\rho = 0.6$ for same traits across domains.
    -   **Conflicts (**$a$ vs $r$): $L = -0.2$ \| $S = -0.5$ \|
        $P = -0.8$ (highest tension).
-   **Group n (Namibia) Specifics:**
    -   **Trait Shift:** Autonomy (a) -0.5; Relatedness (r) +0.5.
    -   **Response Bias:** Added systematic **+0.6 "optimism bias"**.
    -   **Group g (Germany) Specifics:**
    -   **Trait Shift:** Autonomy (a) +0.5; Relatedness (r) -0.5.

## Simulation via `brms`

::: panel-tabset
### Simulated data example

```{r}
#| echo: false
#| message: false
#| warning: false

read_csv("data/simdata.csv") %>%
  slice_sample(n = 6) %>%
  arrange(ID, Domain) %>%
  kable(align = "c")
```

### Base model

```{r}
#| echo: true
#| eval: true

fit <- brm(
  formula = resp ~ 1 + comm + (0 + dim_a + dim_r | ID) + (1 | item),
  data = simdata,
  family = brmsfamily("acat", "logit"), 
  prior = c(
    prior(normal(0, 3), class = "b"), 
    prior(normal(0, 1), class = "sd"),
    prior(lkj(2), class = "cor")
  ),
  backend = "cmdstanr", chains = 4, iter = 2000, warmup = 1000, cores = 4,
  file = "models/fit_rsm"
)

```

### Model summary

```{r}
summary (fit)
```
:::

## Visualise the base model

<style>
.quarto-figure figcaption {
  margin-top: -1.5em;
  line-height: 1.2;
}
</style>

::: panel-tabset
### Wright map

![**Figure 1: Wright map.** Results from the Bayesian 1PL RSM. The
leftmost panel displays item difficulty estimates across the three
domains (L, S, P). The two rightmost panels highlight the
community-level trait
differences.](figures/fig_rsm_recovery.png){width="75%"
fig-align="center"}

### Group Comparison

![**Figure 2: Community Effects.** Conditional effects of Community. The
left plot shows the predicted mean response. The right plot displays
category-wise probabilities (1–4), demonstrating how the **+0.6 optimism
bias** in group $N$ shifts responses toward higher
categories.](figures/fig_comm_effects.png){width="75%"
fig-align="center"}
:::

## Advance model specification

::: panel-tabset
### Advance model

```{r}
#| echo: true
#| eval: true

fit_sep <- brm(
 formula = resp ~ 1 + comm + (0 + Domain:dim_a + Domain:dim_r | ID) + (1 | item),
  data = simdata,
  family = brmsfamily("acat", "logit"), 
  prior = c(
    prior(normal(0, 1), class = "sd"),
    prior(lkj(2), class = "cor")
  ),
  chains = 4, iter = 2000, warmup = 1000, cores = 4, backend = "cmdstanr",
  file = "models/fit_sep_dim"
)
```

### Model summary

```{r}
summary(fit_sep)

```
:::

## Visualise advance model

<style>
.quarto-figure figcaption {
  margin-top: -1.5em;
  line-height: 1.2;
}
</style>

::: panel-tabset
### Domain Correlations

![**Figure 3: Domain-specific latent correlations.** Scatter plots
illustrating the relationship between Autonomy ($a$) and Relatedness
($r$) across Linguistics (L), Social (S), and Professional (P) domains.
Dashed lines represent the linear regressions of the recovered core
conflicts, while point shapes distinguish between the German ($G$) and
Namibian ($N$) communities.](figures/fig_sep_scatter.png){width="75%"
fig-align="center"}

### Correlation Matrix

![**Figure 4: Latent correlation matrix (**$\Phi$) recovery. Heatmaps
comparing the posterior mean estimates (left) against the target truth
values (right) for the $6 \times 6$ latent structure. The model
effectively recovers the simulated
correlations.](figures/fig_phi_heatmap.png){width="75%"
fig-align="center"}
:::

## References

::: {#refs}
:::
